---
title: "Assignment Week 1"
author: "Kat"
output: 
  html_document:
    keep_md: true
---

>
## Step 1: Load the dataset into R & verify data
```{r}
library(UsingR)
data(iris)
str(iris)
summary(iris)
head(iris)
```

>## Step 2: Convert Iris to a data table

```{r}
#install.packages('data.table')
library(data.table)
DT <- as.data.table(iris)
print(DT)
```

>
## Step 3: Begin Outputting Summary Functions
```{r}
DT[,mean(Sepal.Length), by=Species]

DT[,mean(Sepal.Length), by=substring(Species,1,1)]

DT[,.N, by=10*round(Sepal.Length*Sepal.Width/10)]

DT[,.N, by=10*round(Sepal.Length*Sepal.Width/10)]

DT[,.(Count=.N), by=.(Area=10*round(Sepal.Length*Sepal.Width/10))]

```

>
### Step 4: Chaining
```{r}
DT[,.(Sepal_length=median(Sepal.Length),
      Sepal_width=median(Sepal.Width),
      Petal_length=median(Petal.Length),
      Petal_width=median(Petal.Width)),
   by=Species][order(-Species)]

```

> 
### Step 5:Subset of the data table

```{r}
DT[,lapply(.SD,mean),by=Species]
DT[,lapply(.SD,max), by=Species]
DT[,lapply(.SD, max), by=Species]

```

>
### Step 6: Additional Commands

```{r}
iris=as.data.table(iris)
DT1 <- iris[Species == "virginica"]
head(DT1)
DT2 <- iris[Species %in% c("viginica", "versicolor")]
setnames(iris, gsub("^Sepal[.]", "", names(iris)))
iris <- iris[, grep("^Petal", names(iris)) := NULL]
head(iris)

iris[Width*Length>20]
```
>
## I will look at US Cereals and compare nutritional information
```{r}
library(MASS) 
data(UScereal)
head(UScereal)
summary(UScereal)
str(UScereal)
data <- UScereal

```

>
## Loading and looking at the data

```{r}
#Link to explanation of variables
# https://cran.r-project.org/web/packages/MASS/MASS.pdf
library(data.table)
setorder(data, -sugars)
head(data[,c(1,8)],n=10)
```


>
## What are the mean values of across all nutrional content by Manufacturer? Which cereal manufacturer makes the least nutritious cereals overall?

#### Post has the most calories and sugar content in their cereals. It is also slightly higher in fat compared to the others as well as has the 2nd highest sodium content and 2nd highest carb content (24.28 compared with 24.45 for first).

```{r}
#G=General Mills, K=Kelloggs, N=Nabisco, P=Post, Q=Quaker Oats, R=Ralston Purina.

numData <- data[,1:9]
numData <- as.data.table(numData)
sumMFR <- numData[, lapply(.SD, mean), by=mfr]
setorder(sumMFR, -calories, -sugars)
sumMFR

```

>
## What is the mean/median sugar content by manufacturer?
##### It looks like Post has the highest average sugar content in their cereals with 12.7 grams. Interestingly, the median sugar content goes up by about 2 grams when looking at the median, which makes me think they have at least 1 low sugar content cereal bringing their average down. 
```{r}

data <- as.data.table(data)
meanD <- setorder((data[, .(mean=mean(sugars)), by=mfr]), -mean)
meanD
medianD <- setorder((data[, .(median=median(sugars)), by=mfr]), -median)
medianD

```

>
## Which cereal manufacturer makes the relatively healthier cereal based on high protein, high fibre, low calories?

##### It seems that Nabisco makes the cereals with the highest average protein, highest fiber (13.5 g vs 5.35 for next highest), but on the higher side of calories. Sodium content is lowest among these cereals. Fat is second lowest (1.01 g vg .59 g for lowest, avg=1). Sugar content is on the lower than the average (6.606g vs 10.05 avg)
```{r}
setorder(sumMFR, -protein, -fibre, -calories)
sumMFR
```


