---
title: "Data Analytics: Lab Week 1"
author: "Kat Bardash, 10.27.17"
output: 
  html_document:
    keep_md: true
---
####Note: Most exercises were done in 2 different ways, one using data.table, the other with either native R packages or with dplyr.
>
#### First I Will Load in Libraries
```{r}
library(data.table)
library(dplyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```
> 
### Load Directly From a URL

>
## a. Read the csv file from the URL. Paste the command here.
```{r, cache=TRUE}
#data.table way of reading - allows urls

data <- fread("https://www.unsw.adfa.edu.au/australian-centre-for-cyber-security/cybersecurity/ADFA-NB15-Datasets/a%20part%20of%20training%20and%20testing%20set/UNSW_NB15_training-set.csv")

#native R

#data <- as.data.table(read.csv(url("https://www.unsw.adfa.edu.au/australian-centre-for-cyber-security/cybersecurity/ADFA-NB15-Datasets/a%20part%20of%20training%20and%20testing%20set/UNSW_NB15_training-set.csv")))



```

>
##b. Paste the internal structure of your object. 

```{r}
head(data)
str(data)
summary(data)

```

>
##2.a How many tcp, udp, sctp, and arp packets are there in the data set? 
##b. Count occurances of all 131 unique members of this field.

```{r}
#data.table method to get list of counts of all packet types
data2 <- as.data.table(data)

npak <- data2[, .('Number of Rows' =.N), by=proto]
npak

#to see all 132 rows
# print(npak, nrow=132)

# data.table get specific packets
npakS <- data2[proto=="tcp" | proto=="udp" | proto=="sctp" | proto=="arp" , .('Number of Rows' =.N), by=proto]
npakS

# using %in% to get ALL data within those packets
npakS <- subset(data2, proto %in% c('tcp', "udp", "sctp", "arp"))
npakS


# dplyr method
data_df <- as.data.frame(data)
data %>% filter(proto=="tcp" | proto=="udp" | proto=="sctp" | proto=="arp") %>% group_by(proto) %>% summarise(n=n())


```

>##3. What is the mean packet size by protocol? Use the spkts variable. Only provide the first five from your console output.

```{r}

#data.table method
datMean <- data2[, .(mean=mean(spkts)), by=proto]

datMean[1:5,]


#data.table method, using lapply
data2[, lapply(.SD, mean), by=proto, .SDcols=c("spkts")]


# dplyr method
mean_dt <- data %>% group_by(proto) %>% summarize(m=mean(spkts))
head(mean_dt, n=5)


```

>
##4 Five protocols with the greatest mean packet size - descending order 

```{r}
#data.table method
setorder(datMean, -mean)
datMean[1:5,]


#dplyr method
sorted <- arrange(mean_dt, desc(m))
sorted[1:5,]

```

>
## 5 Create a .txt or .csv with provided data and use read.csv(() or read.table to read your file into an object called "my_data". 

```{r}
data5<- read.table("./my_data.txt", sep=",", header=FALSE, nrow=3)
head(data5)
```

>
## b &c. Write an apply() function to calculate sums by row, then by colum
```{r}
#by row
apply(data5, 1, mean)

#by column
apply(data5, 2, mean)

```

>
## d. Write an apply function to calculate sums by row and column. **** Did not *****Use margin=c(1,2)*****
```{r}

rowSum <- apply(data5, 1, sum)
colSum <- apply(data5, 2, mean)

print(rbind(cbind(data5, Rtot=rowSum), Ctot=c(colSum, sum(colSum))))

#  adding 1 to each  ## you need c(1,2) for the WHOLE matrix
apply(data5, c(1,2), function(x) x+1)



```
